---
title: "Plotting Functions and Options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting Functions and Options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

dom <- readRDS("../data/pbmc3k_domino.rds")
```

```{r setup, include = FALSE}
library(domino2)
```

# Visualizing Domino2 data
In addition to the numerous plotting options available in the original [Domino](https://github.com/Elisseeff-Lab/domino), Domino2 has added more functionality and new methods to improve visualization and interpretation of analysis results. Here, we will go over the different plotting functions available, as well as different options that can be utilized, and options for customization. We will be using the domino object from our example analysis on [Getting Started](vignette("domino2")).

## Heatmaps

* `cor_heatmap()`
  * Heatmap can show actual score or be boolean (pick a threshold)
  * All features vs just a subset
  * All receptors, or just some
  * mark_connections displays x when there is a connection
  * Can pass other parameters to `NMF::aheatmap()`

* `feat_heatmap()`
  * normalization
  * min/max thresholding

* `incoming_signaling_heatmap()`
  * Signaling coming into cluster
  * Select clusters
  * min/max Threshold
  * Normalizing

* `signaling_heatmap()`
  * Select clusters if desired
  * Scaling if desired
  * Normalization if desired

## Network Plots

* `gene_network()`
  * Outgoing signaling argument (new!)
  * Can select genes for colors as well
  * Ligand scaling
  * Layout option

* `signaling_network()`
  * Pick outgoing signal
  * Pick incoming signal
  * scaling
  * layout

## Other Types of Plot

* `circos_ligand_receptor()`
  * Circos plots are a new addition!
  * Use cell\_colors argument to change cell colors to match other plots
  * Ligand\_expression\_threshold helps select what signaling to show
  * Can show all clusters or subset to just a few

* `cor_scatter()`
  * Scatter plot for TF vs receptor
  * Match receptor dropout parameter