---
title: "Plotting Functions and Options"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting Functions and Options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE # Just for now since build fails
)
```

```{r setup, include = FALSE}
library(domino2)
```

# Visualizing Domino2 data
In addition to the numerous plotting options available in the original [Domino](https://github.com/Elisseeff-Lab/domino), Domino2 has added more functionality and new methods to improve visualization and interpretation of analysis results. Here, we will go over the different plotting functions available, as well as different options that can be utilized, and options for customization. We will be using the domino object from our example analysis on [Getting Started](vignette("domino2")).

## Updates to Original Plotting Functions:

Functions:
- circos_ligand_receptor
- cor_heatmap
- cor_scatter
- feat_heatmap
- gene_network
- incoming_signaling_heatmap
- signaling_heatmap
- signaling_network

Jacob's stuff I'm cutting from his original vignette:
```{r}
# bug in plotting parameters 
## Error in plot.window(...) : need finite 'xlim' values
# signaling_network(pbmc_dom, edge_weight = .5, max_thresh = 2.5,
#                   showIncomingSignalingClusts = "dendritic_cell")
signaling_network(pbmc_dom, edge_weight = .5, max_thresh = 2.5,
                  showOutgoingSignalingClusts = "CD14_monocyte")
```

The circos plot also includes functionality to change colors used for each cell type to correspond with other figures from the primary analysis of the scRNA-seq data set.

```{r}
cell_types <- unique(pbmc_dom@clusters)
cell_palette <- dittoSeq::dittoColors()[1:length(cell_types)]
names(cell_palette) <- cell_types

circos_ligand_receptor(pbmc_dom, receptor = "CD74", cell_colors = cell_palette)
```